---
title: "Data-Imputation"
author: "Austin Gongora"
date: "4/30/2020"
output: html_document
---



#Libraries
```{r, echo=FALSE}
#The library Mice is used for data imputation 
library(mice)
library(ggplot2)
library(Rtsne)
library(data.table)
library(optparse)
library(RColorBrewer)
library(stringi)
library(caret)
library(ggplot2)
library(RXKCD)
library(tm)
library(wordcloud)
library(RColorBrewer)
library(dplyr)
```



#Data
Using the NBA data set. I will only take the first 100 observations for the sake of the assignments purpose. There are originally over 16,000 observations with 53 columns. 
```{r,echo=FALSE}
players<- read.csv("~/Desktop/NBA/stats/Joined Player Stats.csv")

players <- players [1:5000,]
```

```{r, echo=FALSE}
#Check for missing Values in Players Data Set
anyNA(players)

#Counting Number of NA for each variable 

#Produe dataframe of missing NA values 
x = as.list(sapply(players, function(x) sum(is.na(x))))
z = as.data.frame(x[which(x>0)])

```
The report shows us that there are NA values in the dataset. So we look to do data imputation methods. 


#Example for the assignment. 

We will analyze Variable X3P. Prior and Post imputation
X3P. is a players 3 Point Shot Percentage 
```{r}
cat('mean of X3P. is:' ,mean(players$X3P., na.rm = TRUE))
cat("\nThe median of X3P. is:",median(players$X3P., na.rm = TRUE))
```

#The following figure represents the average 3 Point Shot Percentage by position prior to imputation 
```{r}
p <- ggplot(players, aes(y=X3P., x = Pos), )+ geom_density() +
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1)) + ggtitle("Plot of Avg. 3Point % by Position Pre-Imputation, n = 5000") +
  xlab("Position") + ylab("Shot %")

```
There are multiple variables that contain NA values. We use data imputation to fix this. For instance, there is 2,654 values missing for X3P. feature 

Data Imputation begins.
```{r, echo=FALSE}
init = mice(players, maxit = 0)
meth = init$method
predM = init$predictorMatrix
```


 
#For purposes to show that the imputation is working 
```{r, echo= FALSE}
imputed = mice(players, predictorMatrix = predM, m =2)

perm_impu = imputed

#Imputation Functions
imputed = complete(imputed)
sapply(imputed, function(x) sum(is.na(x)))


anyNA(imputed)
```
Post Imputation there are no longer an NA values in the new dataframe called imputed. We will now perform analysis to test how the two datasets are different. 

```{r}
cat('mean of X3P. is:' ,mean(imputed$X3P., na.rm = TRUE))
cat("\nThe median of X3P. is:",median(imputed$X3P., na.rm = TRUE))
```
Reminder: 
mean of X3P. pre-imputation was: 0.2706842
The median of X3P. pre-imputation was: 0.261



```{r}
q <- ggplot(imputed, aes(y=X3P., x = Pos), )+ 
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1)) + ggtitle("Plot of Avg. 3Point % by Position Post-Imputation, n = 5000") +
  xlab("Position") + ylab("Shot %")


p <- ggplot(players, aes(y=X3P., x = Pos), )+
  stat_summary(fun=mean,geom="line",lwd=2,aes(group=1)) + ggtitle("Plot of Avg. 3Point % by Position Pre-Imputation, n = 5000") +
  xlab("Position") + ylab("Shot %")

```


```{r}
#Feature Engineer new binary variables 
#Yes/No All NBA
#Over Under for Average NBA PLayer POints 

test_imput <- imputed %>%
  mutate(NBA_Binary = ifelse(All.NBA.Selections > 0,1,0)) %>%
  mutate(avg_points = ifelse(PTS > mean(imputed$PTS), 1,0)) %>%
  mutate(good_def = ifelse (BLK. > mean(imputed$BLK.) || STL. > mean(imputed$STL.), 1,0)) %>%
  mutate(ws_hi = ifelse(WS > mean(WS) & (WS >mean(DWS) || (WS > mean(OWS)) ), 1 ,0) ) 
  

#Set factors as numeric variables
#test_imput$Pos<- as.numeric(factor(imputed$Pos))

set.seed("1234")
```

Subset data by player 

1 = Center 
2 = PF
3 = PG
4 = SG 
5 = SF
```{r}
#Postion 1 
new_df <- test_imput%>%
  filter(test_imput$Pos == 4 )%>%
  mutate(efficient = ifelse(PER > mean(PER) ,1,0))

```


```{r}

#TSNE on IMPUTED DATASET
data = subset(new_df , select = -c(X,Player,Tm, Year, All.NBA.Selections, NBA_Binary, Age, G, ws_hi, good_def, avg_points, Pos, X
))


#Create a cosine similarity matrix 
#send Drew plot and explain what is happening 

# read in data (Embeddings & Sentences)

tsne <- Rtsne(data,perplexity =9, stop_lying_iter =195)
sent_tsne<-data.frame(tsne$Y)
setnames(sent_tsne,c("X1","X2"),c("tsne_1","tsne_2"))
#Add labels for color coding purposes
#new_col <- c(rep("IST",5), rep("SRA",2))

#Make new datatable w/ col names 
#new_tsne <- cbind(sent_tsne, Course_Title, Course_credits)
new_tsne <- cbind(sent_tsne, new_df$All.NBA.Selection, new_df$NBA_Binary, new_df$ws_hi, new_df$avg_points, new_df$Pos)

q <- ggplot(new_tsne, aes(x=tsne_1,y=tsne_2,color=new_df$efficient)) +
  geom_point() +
  #theme(legend.position = c(0.3, 0.95)) +
  ggtitle("Colored by Player Efficiency")
q

```