---
title: "Model"
author: "Sameer Sapre"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

We want to estimate the number of All-NBA Selections a player has remaining in their career based on the state of their career. In other words, given their age, and their most recent performances, how many All-NBA selections are they expected to win from now, until the end of their career. 

There is one issue, however, and that is with the model that we would like to use. To estimate the number of All-NBA selections a player will win in the remainder of his career, we would like to use a

```{r}

set.seed(42)
library(tidyverse)
library("car")
library(caret)
df = read.csv('Model Set.csv')
draft = read.csv('Draft.csv') %>% select(Player,Pick)
draft$Pick = as.numeric(draft$Pick)
draft$Player = as.character(draft$Player)
df = left_join(df,draft, by = 'Player')
### Some players are undrafterd and some are drafted in years with more than two rounds,
# I'm adjusting the draft picks below
# Draft Pick #61 --> undrafted
df$Pick[is.na(df$Pick)] <- 61
df$Pick[df$Pick > 61] <- 61
df$Exp[df$Player=="Jaren Jackson" & df$Tm == "MEM"] = 1
# More cleaning for player of interest
df$Player[df$Player=="Jaren Jackson" & df$Tm == "MEM"] = "Jaren Jackson Jr."

```


```{r}
#Make a new variable --> ALL NBA Selections earned thus far
df = df %>% mutate(thusFar = Cumulative.AllNBA - 1)
counts = read.csv('Yearly Article Mentions.csv')
df = left_join(df,counts,by = c("Player","Year"))
df$Mentions[is.na(df$Mentions)] = 0
## Now add the cummulative ESPN mentions
df = df %>% group_by(Player) %>% mutate(ESPN=cumsum(Mentions))
df %>% arrange(desc(ESPN))
```


Keep only retired players
```{r}
latest_year = unique(df$Year)[length(unique(df$Year))]
unretired_players = df %>% filter(Year == latest_year) %>% select(Player) %>% unlist(use.names = F)
df_retired = df %>% filter(!Player %in% unretired_players)
```


From retired players, Find the average career length of "elite" players. Out of the players that received at least 1 All-NBA Selection. The average/median number of years a player plays is 13.

```{r}
df_retired %>% filter(All.NBA.Selections >= 1,) %>% group_by(Player) %>% summarise(n = n(), selections = mean(All.NBA.Selections)) %>% filter(n >= selections) -> usePlayers 
usePlayers %>% select(n) %>% unlist(use.names = F) -> num_years

mean(num_years)
median(num_years)
sample(num_years,size = 100000,replace = T) %>% mean()
```

Only take the first 12 years of retired player's careers - For our players in question, how many All-NBA teams can be expect them to make. By restricting this time interval to a certain number of years, we can meet the assumption of a poisson regression for fixed time intervals.
```{r}
df_retired = df_retired %>% filter(Year != 2019)
```


Lets see if the assumptions for Poisson Regression are met, ie) is mean = variance
```{r}
df_retired$AllNBA.Left %>% mean()
var(df_retired$AllNBA.Left)
```
Looks pretty close

```{r}
## Correlation Matrix
df_retired %>% ungroup %>% select_if(is.numeric) %>% cor() -> corMat
corMat["AllNBA.Left",] %>% abs() %>% sort(decreasing = T)
```


Let's start our regression, but only with retired players
```{r}
#Traditional Stats Model

summary(mod1<- glm(AllNBA.Left ~ Age + FTr + PTS + AST + TRB + X3P + G + Pos + BPM + BLK, family= poisson, data=df_retired))
summary(mod2 <- glm(AllNBA.Left ~ thusFar + ESPN + Age + FTr + AST + PTS + TRB,data = df_retired, family = poisson))
summary(mod3 <- glm(AllNBA.Left ~ WS + PTS + TOV + PER + GS + Pos + STL + AllNBA + Age + FT.,data = df_retired, family = poisson))
```


Cross-Validation by Permutation
```{r eval=FALSE, include=FALSE}

n = nrow(df_retired)
cv_1 = numeric(n)
cv_2 = numeric(n)
for (i in 1:n){
  glm_1 = glm(AllNBA.Left ~ Age + FTr + PTS + AST + TRB + X3P + G + Pos + BPM + BLK, family=poisson, data=df_retired[-i,])
  glm_2 = glm(AllNBA.Left ~ thusFar + ESPN + Age + FTr + AST + PTS + TRB, data = df_retired[-i,], family = poisson)

  new_data = df_retired[i,]

  pred1 = predict(glm_1,new_data,type = "response")
  pred2 = predict(glm_2,new_data, type = "response")
  
  cv_1[i] = (df_retired$AllNBA.Left - pred1)^2
  cv_2[i] = (df_retired$AllNBA.Left - pred2)^2
}
hist(cv_1)
hist(cv_2)
mean(cv_1)
mean(cv_2)
```



What about modeling the chances of a player making a team based on their recent performances?

Let's look at this year's rookies
```{r}
y2019 = df %>% filter(Year == 2019)
preds = predict(mod1,y2019,"response")
y2019$preds = preds
y2019 %>% filter(Exp == 1,Pick <= 5) %>% select(Player,Pick,preds) %>% arrange(desc(preds)) -> rookies2019

rookies2019


```

Let's simulate the careers of the top 5 draft picks from 2019 and graph them
```{r}
plot_df = data.frame()
for (i in seq(1:nrow(rookies2019))){
  dist = rpois(1000,rookies2019$preds[i])
  Player = rep(rookies2019$Player[i],length(dist))
  dist_df = data.frame(Player,dist)
  
  if (nrow(plot_df) == 0){
    plot_df = dist_df
  }
  else {
    plot_df = bind_rows(plot_df,dist_df)
  }
  }
ggplot(plot_df, aes(dist, fill = Player, colour = Player)) +
  geom_density(alpha = 0.3,adjust = 7) + labs(title = "Estimated Amount of All-NBA Selections for 2019 Draft Picks",x = "All-NBA Selections")



```


```{r}
y2019 %>% arrange(desc(preds))
df_retired$fitted = mod1$fitted.values
df_retired %>% arrange(desc(fitted))
```

```{r}
giannis = y2019 %>% filter(Player == "Giannis Antetokounmpo")
a24 = df_retired %>% filter(Age == 24) %>%  arrange(desc(fitted))

a24[1:5,] %>% ggplot(aes(x = Player, y = fitted, fill = Player)) + geom_col() + geom_hline(yintercept = giannis$preds, linetype = "dashed",color = "red") + labs(title = "Giannis vs. All-Time Greats at Age 24", y = "Expected Number of All-NBA Selections Left in Career", caption = "*Red dashed line denotes the expected Number of All-NBA Selections for Giannis Antetokounmpo")
```


```{r}
top10 = y2019 %>% arrange(desc(preds))

top10[1:5,] %>% ggplot(aes(x = Player, y = preds, fill = Player)) + geom_col() + labs(title = "Highest Estimated Future Achievement in the NBA", y = "Expected Number of All-NBA Selections ") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
y2019 %>% ggplot(aes(x = Pos, y = preds, fill = Pos)) + geom_col() + labs(title = "Career Accomplishments by Position", y = "Expected Number of All-NBA Selections ")
```

